---
title: "R Notebook"
output:
  html_document:
    df_print: paged
name: Shiva Neelakantan
student_id: '616916317'
---

# Problem 1
```{r}
patient = read.table('./patient.txt')
names(patient) = c('satisf', 'age', 'severe', 'anxiety')

```
## 1a
```{r}
model1 = lm(satisf~., patient)
summary(model1)
```

## 1b
```{r}
rss = sum(resid(model1)**2)
```
RSS of model1 = 4248.841

## 1c
```{r}
nullmodel = lm(satisf~1, patient)
nullmodel$fitted.values
nullrss = sum(resid(nullmodel)**2)
```
The null model's RSS is larger than the multiple linear regression model, because each value of satisfaction is being predicted by a constant, `1`. Hence, all the null model's fitted values will just be the mean value of satisf in the patient data frame. This causes much higher residuals, on average, than model1. 

## 1d
```{r}
summary(model1)
```

H0 (null): B0=B1=B2=B3=0
H1 (alternative): There exists Bj such that Bj != 0.

F-statistic: 30.05
p-value: 1.542e-10
Assume a significance level (alpha) of 0.05

Our p-value is significantly lower than our established significance level of 0.05. Thus, we can reject the null hypothesis; we have enough evidence that there is a roughly linear relationship between satisfaction and at least one of the predictors.

## 1e
Here, I will choose age as the predictor. 
```{r}
model.e = lm(satisf~age, patient)
model.e
summary(model.e)
```

Let B1 represent the age predictor's coefficient. 
H0 (null): B1 = 0
H1 (alternative): B1 != 0

F-statistic: 71.48
p-value: 9.058e-11
Assume a significance level (alpha) of 0.05

Our p-value is significantly lower than our established significance level of 0.05. Thus, we can reject the null hypothesis; we have enough evidence that there is a roughly linear relationship between one's satisfaction and his or her age.

## 1f
```{r}
pred = predict(model1, data.frame(age=77, severe=68, anxiety=3), interval='prediction', level=0.95)
pred
```
This interval offers the uncertainty from both the reducible and irreducible error of our model's predictions of several data points as well as a single value. This is not the best estimate, for a couple of reasons. First, the range contains a negative value, which is impossible in the real world. Second, the interval is extremey wide, making it hard to draw meaningful conclusions from it.

## 1g
```{r}
diff = predict(lm(satisf~., patient), patient)
```
================================================================================

# Problem 2

## 2a.
The significance level of 0.05 refers to the probability of concluding that there is a statistical difference between, for instance, the results of a placebo drug and an actual drug, even if there is no difference. In other words, we are saying there is a 5% chance that we make the wrong conclusion, due to randomness. 

## 2b.
Although technically, the scientist is correct in that our anxiety predictor does not make a significant difference in our model, due to the p-value of 0.0647, I believe that more data should be gathered and tested with a new model. This is because the p-value is extremely close to our significance level, is a bit ambiguous. Given that the alpha we choose is arbitrary, more testing should occur to ensure that the p-value is indeed greater than 0.05, and is not just an outlier. For instance, if we were to choose a more generous significance level, such as 0.10, the anxiety predictor would pass as statistically significant. 

## 2c.
The scientist's approach seems very naïve and problematic. First of all, there is a chance he could make a mistake in each of the individual tests. For instance, if he carries out 12 individual tests at alpha of 0.10, there is a high probability that we see at least one of the predictors being significant just by chance. 

P(at least one significant predictor) 
  = 1 - P(no significant predictors)
  = 1 - (1 - 0.1) ** 12
  = 1 - (0.9) ** 12
  = 0.72
  
Thus, there is a 72% chance that the scientist concludes that there is at least one significant predictor, without us even knowing any of the T-statistics or p-values. Therefore, his approach may not be the best for his goal of determining if one of the 12 predictors is significant. 

================================================================================

# Problem 3


```{r}
library(ISLR2)
```
## 3a.
```{r}
model.carseats = lm(Sales~., Carseats)
summary(model.carseats)
```

## 3b.

H0: B1==B2==...==Bp==0, where B1, B2, ..., Bp are all the predictors.
H1: There exists one predictor Bk such that Bk != 0

Test Statistic: 
```{r}
model.null = lm(Sales~1, Carseats)
rssR = sum(resid(model.null)**2)
rssF = sum(resid(model.carseats)**2)
dfR = 400-1
dfF = 400-(11+1)
Fstar = ((rssR - rssF)/(dfR-dfF)) / (rssF / dfF)
Fstar
```
Null Distribution: when e ~ N(0, sigma ** 2) and we assume H0, F* has a null distribution of F_(11, 388).
F-statistic: 243.37
p-value from given output: 2.2e-16

Conclusion: At a significance level of 0.05, our p-value indicates that we can reject the null hypothesis. We have enough statistical evidence that at least one of the predictors can predict Sales.

## 3c.
```{r}
model.price = lm(Sales~Price, Carseats)
summary(model.price)
```
H0: Bp == 0
H1: Bp != 0
Null Distribution: when e ~ N(0, sigma ** 2) and we assume H0, F* has a null distribution of F_(11, 388).
F-statistsic: 98.25
p-value: 2.2e-16

Conclusion: At a significance level of 0.05, our p-value indicates that we can reject the null hypothesis. We have enough statistical evidence that Price can be a predictor of Sales. 

# 3d.
According to the model relating Sales to Price, Standard Error = 0.6328. Standard Deviation is just SE * sqrt(n). Thus,

sigma ** 2
  = (0.6328 * sqrt(400)) ** 2
  = (12.656) ** 2
  = 160.174

# 3e.
The fitted model relating Sales to Price has an R^2  of 0.198. This scaled, unit-less value represents the variability in Sales represented by our model. In other words, a higher R^2 implies more variability from the best-fit line, while an R^2 closer to 0 represents minimal variability from the best-fit line. Our model has a rather low R^2, of only 0.198. This signifies a strong, linear correlation between Sales and Price. This conclusion is also reflected by a plot relating the two columns, as shown below. 

```{r}
library(ggplot2)
ggplot(Carseats, aes(Price, Sales)) + geom_point()
```

# 3f.
```{r}
model.shelving = lm(Sales~ShelveLoc, Carseats)
summary(model.shelving)
```
This model shows us the regression coefficients associated with shelving location. Note that this variable is categorical; therefore, ShelveLocGood and ShelveLocBad will not both be included in the coefficients, because one implies the opposite of the other.  ShelveLocGood has a coefficient estimate of roughly 4.69, and ShelveLocMedium has a coefficient estimate of roughly 1.78.

## 3g.
```{r}
carseat.test1 = predict(model.carseats, data.frame(
  CompPrice=mean(Carseats$CompPrice),
  Income=median(Carseats$Income),
  Advertising=15,
  Population=500,
  Price=50,
  ShelveLoc='Good',
  Age=30,
  Education=10,
  Urban='Yes',
  US='Yes'
  ), interval='prediction', level=0.99)
carseat.test1
```
For a confidence level of 0.99, we predict a value of 18.72969, and get a prediction interval of (18.00874, 21.45063). 

## 3h.
```{r}
carseat.test2 = predict(model.carseats, data.frame(
  CompPrice=mean(Carseats$CompPrice),
  Income=median(Carseats$Income),
  Advertising=15,
  Population=500,
  Price=50,
  ShelveLoc='Good',
  Age=30,
  Education=10,
  Urban='Yes',
  US='Yes'
  ), interval='confidence', level=0.99)
carseat.test2
```
For a confidence level of 0.99, we predict a value of 18.72969, and get a confidence interval of (18.06131, 19.39806). 

## 3i.
The two intervals offer the same prediction mean of 18.72969. However, their intervals were different. The confidence interval has a narrower interval than the prediction interval. This is attributed to the fact that a CI does not consider irreducible error, while a PI does. Thus, the PI must offer a wider range to account for irreducible noise in the model.  
