---
title: "CBB_Analysis"
author: "Shiva Neelakantan, Ryan Scehovic, Ryan McNally, Saketh Jonnadula"
date: '2022-05-01'
output: html_document
---


## Year 2013 - Logistic Regression and LDA 
```{r}
df = read.csv('./cbb13.csv')
df[is.na(df)] <- -1
sum(is.na(df))
df$teamYear = paste(df$TEAM, df$YEAR)

library(leaps)
library(ggplot2)
library(VIM)
library(cluster)
library(dplyr)
library(ROCR)
library(MASS)

df = df %>% mutate(WinRate = W/G)
for (i in 1:length(df$POSTSEASON)) {
  if (df$POSTSEASON[i] == 'Champions')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == '2ND')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == 'F4')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == 'E8')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == 'S16')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == 'R32')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == 'R64')
    df$POSTSEASON[i] = 1
  else df$POSTSEASON[i] = 0
}
df$SEED[is.na(df$SEED)] = 0
```



# Q1 SETUP
```{r}
set.seed(7)
df2 <- df[5:22]
df2$WinRate=df$WinRate
df2$POSTSEASON=as.numeric(df2$POSTSEASON)
df2$SEED=df$SEED
train_index = sample(1:nrow(df2),nrow(df2)/4, replace=FALSE)
test_index = -train_index
train = df2[train_index,]
test = df2[-train_index,]
```


# Q1 logistic regression - offense only
```{r}
#ADJOE+ADJDE+EFG_O+EFG_D+TOR+TORD+ORB+DRB+FTR+FTRD+X2P_O+X2P_D+X3P_O+X3P_D+ADJ_T
glm.fit = glm(POSTSEASON~ADJOE+TOR+ORB+FTR+X2P_O+X3P_O+ADJ_T, data=df2, subset=train_index, family='binomial')
summary(glm.fit)
glm.prob = predict(glm.fit, test,type='response') 
glm.pred = rep(0,nrow(test))
glm.pred[glm.prob >0.5] = 1
table(glm.pred,test$POSTSEASON) #rows are predicted, # columns are true 
1-mean(glm.pred == test$POSTSEASON)
cMatrix=table(glm.pred,test$POSTSEASON)
falsenegRate = cMatrix[1,2] / nrow(test)
falsenegRate
falseposRate = cMatrix[2,1] / nrow(test)
falseposRate
ROCRpred <- prediction(glm.prob,test$POSTSEASON)
plot(performance(ROCRpred,'tpr','fpr'))
plot(performance(ROCRpred,'tpr','fpr'),colorize=TRUE,
     print.cutoffs.at=seq(0,1,by=0.05), text.adj=c(-0.2,1.7))
```

# Q1 logistic regression - defense only
```{r}
#ADJOE+ADJDE+EFG_O+EFG_D+TOR+TORD+ORB+DRB+FTR+FTRD+X2P_O+X2P_D+X3P_O+X3P_D+ADJ_T
glm.fit = glm(POSTSEASON~ADJDE+TORD+DRB+FTRD+X3P_D, data=df2, subset=train_index, family='binomial')
summary(glm.fit)
glm.prob = predict(glm.fit, test,type='response') 
glm.pred = rep(0,nrow(test))
glm.pred[glm.prob >0.5] = 1
table(glm.pred,test$POSTSEASON)
1-mean(glm.pred == test$POSTSEASON)
cMatrix=table(glm.pred,test$POSTSEASON)
falsenegRate = cMatrix[1,2] / nrow(test)
falsenegRate
falseposRate = cMatrix[2,1] / nrow(test)
falseposRate
ROCRpred <- prediction(glm.prob,test$POSTSEASON)
plot(performance(ROCRpred,'tpr','fpr'))
plot(performance(ROCRpred,'tpr','fpr'),colorize=TRUE,
     print.cutoffs.at=seq(0,1,by=0.05), text.adj=c(-0.2,1.7))
```


# Q1 LDA - offense only
```{r}
lda.fit = lda(POSTSEASON~ADJOE+EFG_O+TOR+ORB+FTR+X2P_O+X3P_O+ADJ_T, data=df2, subset=train_index)
summary(lda.fit)
lda.pred = predict(lda.fit,test)
table(lda.pred$class,test$POSTSEASON)
mean(lda.pred$class==test$POSTSEASON)
```


# Q1 LDA - defense only
```{r}
lda.fit = lda(POSTSEASON~ADJDE+EFG_D+TORD+DRB+FTRD+X2P_D+X3P_D, data=df2, subset=train_index)
summary(lda.fit)
lda.pred = predict(lda.fit,test)
table(lda.pred$class,test$POSTSEASON)
mean(lda.pred$class==test$POSTSEASON)
```



## Year 2015 - Logistic Regression and LDA 
```{r}
df = read.csv('./cbb15.csv')
df[is.na(df)] <- -1
sum(is.na(df))
df$teamYear = paste(df$TEAM, df$YEAR)

library(leaps)
library(ggplot2)
library(VIM)
library(cluster)
library(dplyr)
library(ROCR)
library(MASS)

df = df %>% mutate(WinRate = W/G)
for (i in 1:length(df$POSTSEASON)) {
  if (df$POSTSEASON[i] == 'Champions')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == '2ND')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == 'F4')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == 'E8')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == 'S16')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == 'R32')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == 'R64')
    df$POSTSEASON[i] = 1
  else df$POSTSEASON[i] = 0
}
df$SEED[is.na(df$SEED)] = 0
```



# Q1 SETUP
```{r}
set.seed(7)
df2 <- df[5:22]
df2$WinRate=df$WinRate
df2$POSTSEASON=as.numeric(df2$POSTSEASON)
df2$SEED=df$SEED
train_index = sample(1:nrow(df2),nrow(df2)/4, replace=FALSE)
test_index = -train_index
train = df2[train_index,]
test = df2[-train_index,]
```


# Q1 logistic regression - offense only
```{r}
#ADJOE+ADJDE+EFG_O+EFG_D+TOR+TORD+ORB+DRB+FTR+FTRD+X2P_O+X2P_D+X3P_O+X3P_D+ADJ_T
glm.fit = glm(POSTSEASON~ADJOE+TOR+ORB+FTR+X2P_O+X3P_O+ADJ_T, data=df2, subset=train_index, family='binomial')
summary(glm.fit)
glm.prob = predict(glm.fit, test,type='response') 
glm.pred = rep(0,nrow(test))
glm.pred[glm.prob >0.5] = 1
table(glm.pred,test$POSTSEASON) #rows are predicted, # columns are true 
1-mean(glm.pred == test$POSTSEASON)
cMatrix=table(glm.pred,test$POSTSEASON)
falsenegRate = cMatrix[1,2] / nrow(test)
falsenegRate
falseposRate = cMatrix[2,1] / nrow(test)
falseposRate
ROCRpred <- prediction(glm.prob,test$POSTSEASON)
plot(performance(ROCRpred,'tpr','fpr'))
plot(performance(ROCRpred,'tpr','fpr'),colorize=TRUE,
     print.cutoffs.at=seq(0,1,by=0.05), text.adj=c(-0.2,1.7))
```

# Q1 logistic regression - defense only
```{r}
#ADJOE+ADJDE+EFG_O+EFG_D+TOR+TORD+ORB+DRB+FTR+FTRD+X2P_O+X2P_D+X3P_O+X3P_D+ADJ_T
glm.fit = glm(POSTSEASON~ADJDE+TORD+DRB+FTRD+X3P_D, data=df2, subset=train_index, family='binomial')
summary(glm.fit)
glm.prob = predict(glm.fit, test,type='response') 
glm.pred = rep(0,nrow(test))
glm.pred[glm.prob >0.5] = 1
table(glm.pred,test$POSTSEASON)
1-mean(glm.pred == test$POSTSEASON)
cMatrix=table(glm.pred,test$POSTSEASON)
falsenegRate = cMatrix[1,2] / nrow(test)
falsenegRate
falseposRate = cMatrix[2,1] / nrow(test)
falseposRate
ROCRpred <- prediction(glm.prob,test$POSTSEASON)
plot(performance(ROCRpred,'tpr','fpr'))
plot(performance(ROCRpred,'tpr','fpr'),colorize=TRUE,
     print.cutoffs.at=seq(0,1,by=0.05), text.adj=c(-0.2,1.7))
```


# Q1 LDA - offense only
```{r}
lda.fit = lda(POSTSEASON~ADJOE+EFG_O+TOR+ORB+FTR+X2P_O+X3P_O+ADJ_T, data=df2, subset=train_index)
summary(lda.fit)
lda.pred = predict(lda.fit,test)
table(lda.pred$class,test$POSTSEASON)
mean(lda.pred$class==test$POSTSEASON)
```


# Q1 LDA - defense only
```{r}
lda.fit = lda(POSTSEASON~ADJDE+EFG_D+TORD+DRB+FTRD+X2P_D+X3P_D, data=df2, subset=train_index)
summary(lda.fit)
lda.pred = predict(lda.fit,test)
table(lda.pred$class,test$POSTSEASON)
mean(lda.pred$class==test$POSTSEASON)
```


## Year 2017 - Logistic Regression and LDA 
```{r}
df = read.csv('./cbb17.csv')
df[is.na(df)] <- -1
sum(is.na(df))
df$teamYear = paste(df$TEAM, df$YEAR)

library(leaps)
library(ggplot2)
library(VIM)
library(cluster)
library(dplyr)
library(ROCR)
library(MASS)

df = df %>% mutate(WinRate = W/G)
for (i in 1:length(df$POSTSEASON)) {
  if (df$POSTSEASON[i] == 'Champions')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == '2ND')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == 'F4')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == 'E8')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == 'S16')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == 'R32')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == 'R64')
    df$POSTSEASON[i] = 1
  else df$POSTSEASON[i] = 0
}
df$SEED[is.na(df$SEED)] = 0
```



# Q1 SETUP
```{r}
set.seed(7)
df2 <- df[5:22]
df2$WinRate=df$WinRate
df2$POSTSEASON=as.numeric(df2$POSTSEASON)
df2$SEED=df$SEED
train_index = sample(1:nrow(df2),nrow(df2)/4, replace=FALSE)
test_index = -train_index
train = df2[train_index,]
test = df2[-train_index,]
```


# Q1 logistic regression - offense only
```{r}
#ADJOE+ADJDE+EFG_O+EFG_D+TOR+TORD+ORB+DRB+FTR+FTRD+X2P_O+X2P_D+X3P_O+X3P_D+ADJ_T
glm.fit = glm(POSTSEASON~ADJOE+TOR+ORB+FTR+X2P_O+X3P_O+ADJ_T, data=df2, subset=train_index, family='binomial')
summary(glm.fit)
glm.prob = predict(glm.fit, test,type='response') 
glm.pred = rep(0,nrow(test))
glm.pred[glm.prob >0.5] = 1
table(glm.pred,test$POSTSEASON) #rows are predicted, # columns are true 
1-mean(glm.pred == test$POSTSEASON)
cMatrix=table(glm.pred,test$POSTSEASON)
falsenegRate = cMatrix[1,2] / nrow(test)
falsenegRate
falseposRate = cMatrix[2,1] / nrow(test)
falseposRate
ROCRpred <- prediction(glm.prob,test$POSTSEASON)
plot(performance(ROCRpred,'tpr','fpr'))
plot(performance(ROCRpred,'tpr','fpr'),colorize=TRUE,
     print.cutoffs.at=seq(0,1,by=0.05), text.adj=c(-0.2,1.7))
```

# Q1 logistic regression - defense only
```{r}
#ADJOE+ADJDE+EFG_O+EFG_D+TOR+TORD+ORB+DRB+FTR+FTRD+X2P_O+X2P_D+X3P_O+X3P_D+ADJ_T
glm.fit = glm(POSTSEASON~ADJDE+TORD+DRB+FTRD+X3P_D, data=df2, subset=train_index, family='binomial')
summary(glm.fit)
glm.prob = predict(glm.fit, test,type='response') 
glm.pred = rep(0,nrow(test))
glm.pred[glm.prob >0.5] = 1
table(glm.pred,test$POSTSEASON)
1-mean(glm.pred == test$POSTSEASON)
cMatrix=table(glm.pred,test$POSTSEASON)
falsenegRate = cMatrix[1,2] / nrow(test)
falsenegRate
falseposRate = cMatrix[2,1] / nrow(test)
falseposRate
ROCRpred <- prediction(glm.prob,test$POSTSEASON)
plot(performance(ROCRpred,'tpr','fpr'))
plot(performance(ROCRpred,'tpr','fpr'),colorize=TRUE,
     print.cutoffs.at=seq(0,1,by=0.05), text.adj=c(-0.2,1.7))
```


# Q1 LDA - offense only
```{r}
lda.fit = lda(POSTSEASON~ADJOE+EFG_O+TOR+ORB+FTR+X2P_O+X3P_O+ADJ_T, data=df2, subset=train_index)
summary(lda.fit)
lda.pred = predict(lda.fit,test)
table(lda.pred$class,test$POSTSEASON)
mean(lda.pred$class==test$POSTSEASON)
```


# Q1 LDA - defense only
```{r}
lda.fit = lda(POSTSEASON~ADJDE+EFG_D+TORD+DRB+FTRD+X2P_D+X3P_D, data=df2, subset=train_index)
summary(lda.fit)
lda.pred = predict(lda.fit,test)
table(lda.pred$class,test$POSTSEASON)
mean(lda.pred$class==test$POSTSEASON)
```





## Year 2019 - Logistic Regression and LDA 
```{r}
df = read.csv('./cbb19.csv')
df[is.na(df)] <- -1
sum(is.na(df))
df$teamYear = paste(df$TEAM, df$YEAR)

library(leaps)
library(ggplot2)
library(VIM)
library(cluster)
library(dplyr)
library(ROCR)
library(MASS)

df = df %>% mutate(WinRate = W/G)
for (i in 1:length(df$POSTSEASON)) {
  if (df$POSTSEASON[i] == 'Champions')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == '2ND')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == 'F4')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == 'E8')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == 'S16')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == 'R32')
    df$POSTSEASON[i] = 1
  else if (df$POSTSEASON[i] == 'R64')
    df$POSTSEASON[i] = 1
  else df$POSTSEASON[i] = 0
}
df$SEED[is.na(df$SEED)] = 0
```



# Q1 SETUP
```{r}
set.seed(7)
df2 <- df[5:22]
df2$WinRate=df$WinRate
df2$POSTSEASON=as.numeric(df2$POSTSEASON)
df2$SEED=df$SEED
train_index = sample(1:nrow(df2),nrow(df2)/4, replace=FALSE)
test_index = -train_index
train = df2[train_index,]
test = df2[-train_index,]
```


# Q1 logistic regression - offense only
```{r}
#ADJOE+ADJDE+EFG_O+EFG_D+TOR+TORD+ORB+DRB+FTR+FTRD+X2P_O+X2P_D+X3P_O+X3P_D+ADJ_T
glm.fit = glm(POSTSEASON~ADJOE+TOR+ORB+FTR+X2P_O+X3P_O+ADJ_T, data=df2, subset=train_index, family='binomial')
summary(glm.fit)
glm.prob = predict(glm.fit, test,type='response') 
glm.pred = rep(0,nrow(test))
glm.pred[glm.prob >0.5] = 1
table(glm.pred,test$POSTSEASON) #rows are predicted, # columns are true 
1-mean(glm.pred == test$POSTSEASON)
cMatrix=table(glm.pred,test$POSTSEASON)
falsenegRate = cMatrix[1,2] / nrow(test)
falsenegRate
falseposRate = cMatrix[2,1] / nrow(test)
falseposRate
ROCRpred <- prediction(glm.prob,test$POSTSEASON)
plot(performance(ROCRpred,'tpr','fpr'))
plot(performance(ROCRpred,'tpr','fpr'),colorize=TRUE,
     print.cutoffs.at=seq(0,1,by=0.05), text.adj=c(-0.2,1.7))
```

# Q1 logistic regression - defense only
```{r}
#ADJOE+ADJDE+EFG_O+EFG_D+TOR+TORD+ORB+DRB+FTR+FTRD+X2P_O+X2P_D+X3P_O+X3P_D+ADJ_T
glm.fit = glm(POSTSEASON~ADJDE+TORD+DRB+FTRD+X3P_D, data=df2, subset=train_index, family='binomial')
summary(glm.fit)
glm.prob = predict(glm.fit, test,type='response') 
glm.pred = rep(0,nrow(test))
glm.pred[glm.prob >0.5] = 1
table(glm.pred,test$POSTSEASON)
1-mean(glm.pred == test$POSTSEASON)
cMatrix=table(glm.pred,test$POSTSEASON)
falsenegRate = cMatrix[1,2] / nrow(test)
falsenegRate
falseposRate = cMatrix[2,1] / nrow(test)
falseposRate
ROCRpred <- prediction(glm.prob,test$POSTSEASON)
plot(performance(ROCRpred,'tpr','fpr'))
plot(performance(ROCRpred,'tpr','fpr'),colorize=TRUE,
     print.cutoffs.at=seq(0,1,by=0.05), text.adj=c(-0.2,1.7))
```


# Q1 LDA - offense only
```{r}
lda.fit = lda(POSTSEASON~ADJOE+EFG_O+TOR+ORB+FTR+X2P_O+X3P_O+ADJ_T, data=df2, subset=train_index)
summary(lda.fit)
lda.pred = predict(lda.fit,test)
table(lda.pred$class,test$POSTSEASON)
mean(lda.pred$class==test$POSTSEASON)
```


# Q1 LDA - defense only
```{r}
lda.fit = lda(POSTSEASON~ADJDE+EFG_D+TORD+DRB+FTRD+X2P_D+X3P_D, data=df2, subset=train_index)
summary(lda.fit)
lda.pred = predict(lda.fit,test)
table(lda.pred$class,test$POSTSEASON)
mean(lda.pred$class==test$POSTSEASON)
```



